---
title: "p8105_hw1_mm5351"
author: Martha Mulugeta
date: 2019-09-14
output: github_document
---

**Problem 1**
```{r hw1_df}
library(tidyverse)
set.seed(1)
hw1_df = tibble(
sample = rnorm(8),
vec_logical = c(sample > 0),
vec_char = c(length = 8),
vec_factor = factor(c(length = 8)), levels = c(3),

mean(sample),
mean(vec_logical),
mean(vec_char),
mean(vec_factor),
)
```
The means of the sample, logical vector and character vector were produced. However, the mean for the factor vector was not applicable. This is likely because factor variables represent categories, and the average of cannot be taken of categories. The logical vector produced "True" or "False" variables coded as 1 and 0 respectively. Therefore, a mean can be taken of those values. The sample and character vector both contained numbers, of which means are easily calculated. 

```{r convert_to_numeric, eval = FALSE}
as.numeric(pull(hw1_df, vec_logical))
as.numeric(pull(hw1_df, vec_char))
as.numeric(pull(hw1_df, vec_factor))
```
When converting variables from the logical, character, and factor vectors into numeric vectors, the means for all are calculable. This is because it a numeral represents each variable, of which the mean can easily be calculated. 

```{r convert, eval = FALSE}
as.numeric(pull(hw1_df, vec_logical)) * sample(pull(hw1_df, sample))
as.factor(pull(hw1_df, vec_logical)) * sample(pull(hw1_df, sample))
as.numeric(as.factor(pull(hw1_df, vec_logical))) * sample(pull(hw1_df, sample))
```
**Problem 2**
```{r plot_df}
set.seed(2)
plot_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  vec_logical = c(x + y > 1),
  vec_numeric = as.numeric(vec_logical), 
  vec_factor = as.factor(vec_logical)
)
```

There are `r nrow(plot_df)` rows and `r ncol(plot_df)` columns in the dataset plot_df. The mean of x is `r mean(pull(plot_df, x))`. The median of x is `r median(pull(plot_df, x))`. The standard deviation of x is `r sd(pull(plot_df, x))`. The proportion of cases for which x + y > 1 is	`r mean(pull(plot_df, vec_numeric))`.  

```{r scatterplot}
ggplot(plot_df, aes(x = x, y = y, color = vec_logical)) + geom_point()
ggplot(plot_df, aes(x = x, y = y, color = vec_numeric)) + geom_point()
ggplot(plot_df, aes(x = x, y = y, color = vec_factor)) + geom_point()
ggsave("scatter_plot.pdf")
```