---
title: "p8105_hw1_mm5351"
author: Martha Mulugeta
date: 2019-09-14
output: github_document
---

**Problem 1:**
*Create a data frame with a random sample of size 8 from a standard Normal distribution, a logical vector indicating whether elements of the sample are greater than 0, a character vector of length 8, and a factor vector of length 8 with 3 different factor “levels”. Take the means of each variable in the dataframe.*
```{r hw1_df}
library(tidyverse)

##Create dataframe
set.seed(1)
hw1_df = tibble(
sample = rnorm(8),
vec_logical = c(sample > 0),
vec_char = c(length = 8),
vec_factor = factor(c(length = 8)), levels = c(3),

##Take means of each variable
mean(sample),
mean(vec_logical),
mean(vec_char),
mean(vec_factor),
)
```
The means of the sample, logical vector and character vector were produced. However, the mean for the factor vector was not applicable. This is likely because factor variables represent categories, and the average of cannot be taken of categories. The logical vector produced "True" or "False" variables coded as 1 and 0 respectively. Therefore, a mean can be taken of those values. The sample and character vector both contained numbers, of which means are easily calculated. 

*Write a code chunk that applies the as.numeric function to the logical, character, and factor variables showing the chunk but not the output.* 
```{r convert_to_numeric, eval = FALSE}
##Convert vectors
as.numeric(pull(hw1_df, vec_logical))
as.numeric(pull(hw1_df, vec_char))
as.numeric(pull(hw1_df, vec_factor))
```
When converting variables from the logical, character, and factor vectors into numeric vectors, the means for all are calculable. This is because a numeral represents each variable, of which the mean can easily be calculated. 

*Convert the logical vector to numeric, and multiply the random sample by the result, convert the logical vector to a factor, and multiply the random sample by the result, and convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result.*
```{r convert}
##Convert and manipulate vectors 
as.numeric(pull(hw1_df, vec_logical)) * sample(pull(hw1_df, sample))
as.factor(pull(hw1_df, vec_logical)) * sample(pull(hw1_df, sample))
as.numeric(as.factor(pull(hw1_df, vec_logical))) * sample(pull(hw1_df, sample))
```

**Problem 2:**
*Create a data frame where x is a random sample of size 500 from a standard Normal distribution and y is a random sample of size 500 from a standard Normal distribution. Create a logical vector indicating whether x + y > 1, a numeric vector created by coercing the above logical vector, and a factor vector created by coercing the above logical vector.* 
```{r plot_df}
##Create dataframe
set.seed(2)
plot_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  vec_logical = c(x + y > 1),
  vec_numeric = as.numeric(vec_logical), 
  vec_factor = as.factor(vec_logical)
)
```

There are `r nrow(plot_df)` rows and `r ncol(plot_df)` columns in the dataset plot_df. The mean of x is `r mean(pull(plot_df, x))`. The median of x is `r median(pull(plot_df, x))`. The standard deviation of x is `r sd(pull(plot_df, x))`. The proportion of cases for which x + y > 1 is	`r mean(pull(plot_df, vec_numeric))`.  

*Make a scatterplot of y vs x; color point using the logical variable. Make a second and third scatterplot that color prints using the numeric and factor variables, respectively.Export the first scatter plot.* 
```{r scatterplot}
##Scatterplot 1: logical variable
ggplot(plot_df, aes(x = x, y = y, color = vec_logical)) + geom_point()

##Scatterplot 2: numeric variable
ggplot(plot_df, aes(x = x, y = y, color = vec_numeric)) + geom_point()

##Scatterplot 3: factor variable
ggplot(plot_df, aes(x = x, y = y, color = vec_factor)) + geom_point()

##Save first scatterplot
ggplot(plot_df, aes(x = x, y = y, color = vec_logical)) + geom_point()
ggsave("scatter_plot.pdf")
```

In the first scatterplot, red represents false (x + y =< 1) and blue represents true (x + y > 1).In the second scatterplot, the lighter blue is indicative of a value 1 (x + y > 1), and the darker blue is indicative of a value of 0 (x + y =< 1). Despite there being a gradient present, the actual values in the dataset are either 1 or 0. In the third scatterplot, red represents false (x + y =< 1) and blue represents true (x + y > 1). This is the same as that of the first scatterplot.